Bipartite Fidelity and Loschmidt Echo of Bosonic Conformal Interface

### 2017-05-07 general_formulation.tex
[draft] 1.add bibs, 2. geneneral formulation, free boson lattice done 3. Appendix of winding mode not done yet, I will add the caveat later

We consider a conformal interface connecting two potentially different CFTs in $1+1$ dimensions. The interface is located at $x=0$ and characterized by the ``gluing condition''
\begin{eqnarray}\begin{aligned}
\label{eq:def_M}
\begin{pmatrix}
\partial_x\phi\\
\partial_t\phi
\end{pmatrix}_{x=0^-}
=M\begin{pmatrix}
\partial_x\phi\\
\partial_t\phi
\end{pmatrix}_{x=0^+}
\end{aligned}\end{eqnarray}
The interface is assumed to be penetrable. Instead of requiring the normal components of the stress tensor to vanish\cite{cardy_conformal_1984}, one only needs the stress tensor to be continuous across the defect. From the explicit expression of the stress tensor $T^{xt}=-\partial_x\phi\partial_t\phi$, one can show that\cite{bachas_permeable_2002} $M$ is an element of the Lorentz group $O(1,1)$. In terms of the boost parametrization, we have
\begin{eqnarray}\begin{aligned}
M_1(\theta)=\pm
\begin{bmatrix}
\lambda^{-1} & 0 \\
0 & \lambda
\end{bmatrix}\quad
M_2(\theta)=\pm
\begin{bmatrix}
0 & \lambda  \\
\lambda^{-1} & 0 
\end{bmatrix}
\end{aligned}\end{eqnarray}
where $\lambda=\tan\theta$ for $\theta\in\left[-\frac{\pi}{2},\frac{\pi}{2}\right]$.

We note the following special cases. For $\theta=0,\pm\pi/2$, $\lambda$ (or $\lambda^{-1}$) appears to be singular and the fields on either side of the defect cannot transmit to the other side. For example, $M_1(0)$ implies $\partial_x\phi(0^+)=\partial_t\phi(0^-)=0$ which correspond to Dirichlet boundary condition on the left, and Neumann condition on the right. Hereafter we shall denote it as `DN'. Similarly one can check that $M_1(\pm\pi/2),M_2(0),M_2(\pm\pi/2)$ correspond to ND, NN, DD respectively. For $\theta=\pm\pi/4$, on the other hand, $M(\theta)$ characterizes a perfectly transmitting defect. For example, there is effectively no defect in the case $M(\pi/4)$. For the other three cases, although the field $\phi$ may pick up a phase across the defect, the two counter propagating waves are still fully transmitted. {\bf\color{red} shall we follow permeable wall paper to denote them as PP, etc? it is a bit confusing though..}

It proves useful to rewrite Eq.~\eqref{eq:def_M} in the ``light cone'' coordinate $x^\pm\equiv t\pm x$, where the physical meaning of $\theta$ is most transparent. For clarity, we denote the field on the negative (positive) real axis as $\phi^1$ ($\phi^2$). {\bf\color{red}Tianci: Add in the right panel in Fig.~\ref{fig:cut-and-join} the field content?}. After some algebra, we have
\begin{eqnarray}\begin{aligned}
\label{eq:def_S}
\begin{pmatrix}
\partial_+\phi^1\\
\partial_-\phi^2
\end{pmatrix}
=S
\begin{pmatrix}
\partial_-\phi^1\\
\partial_+\phi^2
\end{pmatrix}
\end{aligned}\end{eqnarray}
where 
\begin{eqnarray}\begin{aligned}
\label{eq:S1_S2}
S_1(\theta)=\begin{bmatrix}
\cos 2\theta & \sin 2\theta \\
\sin 2\theta & -\cos 2\theta
\end{bmatrix}\quad
S_2(\theta)=\begin{bmatrix}
-\cos 2\theta & \sin 2\theta \\
-\sin 2\theta & -\cos 2\theta
\end{bmatrix}
\end{aligned}\end{eqnarray}
The scattering matrix $S_1$ and $S_2$ relate the incoming and outgoing waves at the defect. For example, $S_1(0)$ implies that $\partial_\pm\phi^{1}=\pm\partial_\mp\phi^{2}$ as discussed above. For other nontrivial values of $\theta$, the defect becomes partially-transmitting. The transmission and reflection coefficients, which are determined by $\theta$, can be read off from the scattering matrices. As we shall show explicitly in Sec.~\ref{sec_sub:free_boson_lattice}, the scattering matrices are independent of the wavelengths, as required by the scale invariance.

The problem of conformal interface can be studied in the framework of boundary CFT\cite{cardy_boundary_2004,cardy_conformal_1984} after we fold the field $\phi^2$ to the left half plane on top of $\phi^1$\cite{oshikawa_boundary_2010}. The boundary at $x=0$ becomes impenetrable for the tensor theory, and there hosts a conformal boundary states for the bosons. The folding sends $\phi^2(x)\rightarrow\phi^2(-x)$ for $x<0$, therefore the gluing condition reads
\begin{eqnarray}\begin{aligned}
\partial_t(\cos\theta\phi^1-\sin\theta\phi^2)=0 \quad
\partial_x(\sin\theta\phi^1+\cos\theta\phi^2)=0 
\end{aligned}\end{eqnarray}
for the case $M=M_1(\theta)$. After performing explicit mode expansion for both $\phi^{1,2}$, {\bf\color{red}write down mode expansion, or simply write down the physical meanings of a's as follows?}we have
\begin{eqnarray}\begin{aligned}
\label{eq:rotation_a_basis}
\cos\theta a_n^1-\sin\theta a_n^2 &= +( \cos\theta\bar{a}_{-n}^1-\sin\theta \bar{a}_{-n}^2 ) \\
\sin\theta a_n^1+\cos\theta a_n^2 &= -( \sin\theta\bar{a}_{-n}^1+\cos\theta \bar{a}_{-n}^2 ) 
\end{aligned}\end{eqnarray}
where $n>0$ is an integer. $a_n^{1,2}$ ($\bar{a}_n^{1,2}$) annihilate the left (right)-moving modes of $\phi^{1,2}$, and $a_{-n}^{1,2}$ ($\bar{a}_{-n}^{1,2}$) are the corresponding creation operators. Eq.~\eqref{eq:rotation_a_basis} is essentially a rotation of basis, and the conformal boundary state reads\cite{oshikawa_boundary_2010}
\begin{equation}
\label{eq:bd_state}
\begin{aligned}
| B \rangle 
& =  \exp\Big\{ \sum_{n > 0 } \frac{1}{n}
\begin{pmatrix}
a_{-n}^1\\
a_{-n}^2\\                              
\end{pmatrix}
S_1
\begin{pmatrix}
\bar{a}_{-n}^1  \bar{a}_{-n}^2
\end{pmatrix} \Big\} |0\rangle
\end{aligned}
\end{equation}
where $S^1$ is precisely the scattering matrix in Eq.~\eqref{eq:def_S}. As one can explicitly check, Eq.~\eqref{eq:bd_state} satisfies the condition for the conformal boundary state\cite{oshikawa_boundary_2010}. The calculation for the case $M=M_2(\theta)$ is completely analogue. Eq.~\eqref{eq:bd_state} will be the starting point of calculating fidelity and Loschmidt echo in Sec.~\ref{sec:analytic_numerics}. In particular, 
\todo[inline]{say sth: connecting to what will follow etc. }

>>>>

We consider a conformal interface connecting two potentially different CFTs in $1+1$ dimensions. The interface is located at $x=0$ and characterized by the ``gluing condition''
\begin{eqnarray}\begin{aligned}
\label{eq:def_M}
\begin{pmatrix}
\partial_x\phi\\
\partial_t\phi
\end{pmatrix}_{x=0^-}
=M\begin{pmatrix}
\partial_x\phi\\
\partial_t\phi
\end{pmatrix}_{x=0^+}
\end{aligned}\end{eqnarray}
The interface is assumed to be penetrable, and the stress tensor is continuous across the defect. From the explicit expression of the stress tensor $T^{xt}=-\partial_x\phi\partial_t\phi$, one can show that\cite{bachas_permeable_2002} $M$ is an element of the Lorentz group $O(1,1)$. In terms of the boost parametrization, we have
\begin{eqnarray}\begin{aligned}
M_1(\theta)=\pm
\begin{bmatrix}
\lambda^{-1} & 0 \\
0 & \lambda
\end{bmatrix}\quad
M_2(\theta)=\pm
\begin{bmatrix}
0 & \lambda  \\
\lambda^{-1} & 0 
\end{bmatrix}
\end{aligned}\end{eqnarray}
where $\lambda=\tan\theta$ for $\theta\in\left[-\frac{\pi}{2},\frac{\pi}{2}\right]$.

We note the following special cases. For $\theta=0,\pm\pi/2$, $\lambda$ (or $\lambda^{-1}$) appears to be singular and the fields on either side of the defect cannot transmit to the other side. For example, $M_1(0)$ implies $\partial_x\phi(0^+)=\partial_t\phi(0^-)=0$ which correspond to Dirichlet boundary condition on the left, and Neumann condition on the right. Hereafter we shall denote it as `DN'. Similarly one can check that $M_1(\pm\pi/2),M_2(0),M_2(\pm\pi/2)$ correspond to ND, NN, DD respectively. For $\theta=\pm\pi/4$, on the other hand, $M(\theta)$ characterizes a perfectly transmitting defect. For example, there is effectively no defect in the case $M(\pi/4)$. For the other three cases, although the field $\phi$ may pick up a phase across the defect, the two counter propagating waves are still fully transmitted.
\todo[inline]{shall we follow permeable wall paper to denote them as PP, etc? it is a bit confusing though..}

It proves useful to rewrite Eq.~\eqref{eq:def_M} in the ``light cone'' coordinate $x^\pm\equiv t\pm x$, where the physical meaning of $\theta$ is most transparent. For clarity, we denote the field on the negative (positive) real axis as $\phi^1$ ($\phi^2$).
\todo[inline]{Tianci: Add in the right panel in Fig.~\ref{fig:cut-and-join} the field content?}
After some algebra, we have
\begin{eqnarray}\begin{aligned}
\label{eq:def_S}
\begin{pmatrix}
\partial_+\phi^1\\
\partial_-\phi^2
\end{pmatrix}
=S
\begin{pmatrix}
\partial_-\phi^1\\
\partial_+\phi^2
\end{pmatrix}
\end{aligned}\end{eqnarray}
where 
\begin{eqnarray}\begin{aligned}
\label{eq:S1_S2}
S_1(\theta)=\begin{bmatrix}
\cos 2\theta & \sin 2\theta \\
\sin 2\theta & -\cos 2\theta
\end{bmatrix}\quad
S_2(\theta)=\begin{bmatrix}
-\cos 2\theta & \sin 2\theta \\
-\sin 2\theta & -\cos 2\theta
\end{bmatrix}
\end{aligned}\end{eqnarray}
The scattering matrix $S_1$ and $S_2$ relate the incoming and outgoing waves at the defect. For example, $S_1(0)$ implies that $\partial_\pm\phi^{1}=\pm\partial_\mp\phi^{2}$ as discussed above. For other nontrivial values of $\theta$, the defect becomes partially-transmitting, and the transmission and reflection coefficients can be read off from the scattering matrices. In particular, these coefficients depend only on the parameter $\theta$ of the defect, not on the wavelength of the incoming wave, which is in contrast to the case in mesoscopic system.\cite{datta_1995} This is expected due to the scale invariance.

The problem of conformal interface can be studied in the framework of boundary CFT\cite{cardy_boundary_2004,cardy_conformal_1984} after we fold the field $\phi^2$ to the left half plane on top of $\phi^1$\cite{oshikawa_boundary_2010}. The boundary at $x=0$ becomes impenetrable for the folded system, and the resulting tensor theory admits a conformal boundary states for the bosons. The folding sends $\phi^2(x)\rightarrow\phi^2(-x)$ for $x<0$, therefore the gluing condition reads
\begin{eqnarray}\begin{aligned}
\partial_t(\cos\theta\phi^1-\sin\theta\phi^2)=0 \quad
\partial_x(\sin\theta\phi^1+\cos\theta\phi^2)=0 
\end{aligned}\end{eqnarray}
for the case $M=M_1(\theta)$. After performing explicit mode expansion for both $\phi^{1,2}$, we have
\begin{eqnarray}\begin{aligned}
\label{eq:rotation_a_basis}
\cos\theta a_n^1-\sin\theta a_n^2 &= +( \cos\theta\bar{a}_{-n}^1-\sin\theta \bar{a}_{-n}^2 ) \\
\sin\theta a_n^1+\cos\theta a_n^2 &= -( \sin\theta\bar{a}_{-n}^1+\cos\theta \bar{a}_{-n}^2 ) 
\end{aligned}\end{eqnarray}
where $n>0$ is an integer. $a_n^{1,2}$ ($\bar{a}_n^{1,2}$) annihilate the left (right)-moving modes of $\phi^{1,2}$, and $a_{-n}^{1,2}$ ($\bar{a}_{-n}^{1,2}$) are the corresponding creation operators. Eq.~\eqref{eq:rotation_a_basis} is essentially a rotation of basis, and the conformal boundary state reads\cite{oshikawa_boundary_2010}
\begin{equation}
\label{eq:bd_state}
\begin{aligned}
| B \rangle 
& =  \exp\Big\{ \sum_{n > 0 } \frac{1}{n}
\begin{pmatrix}
a_{-n}^1\\
a_{-n}^2\\                              
\end{pmatrix}
S_1
\begin{pmatrix}
\bar{a}_{-n}^1  \bar{a}_{-n}^2
\end{pmatrix} \Big\} |0\rangle
\end{aligned}
\end{equation}
where $S^1$ is precisely the scattering matrix in Eq.~\eqref{eq:def_S}. As one can explicitly check, Eq.~\eqref{eq:bd_state} satisfies the condition for the conformal boundary state\cite{oshikawa_boundary_2010}. The calculation for the case $M=M_2(\theta)$ is completely analogue. Eq.~\eqref{eq:bd_state} will be used extensively in the boundary state calculation of fidelity and Loschmidt echo in Sec.~\ref{sec:analytic_numerics}.

\todo[inline]{say sth: connecting to what will follow etc? }

### 2017-05-07 gnd_dn_lambda.tex
[draft] 1.add bibs, 2. geneneral formulation, free boson lattice done 3. Appendix of winding mode not done yet, I will add the caveat later

In this appendix, we would like to calculate the amplitude for the setup shown in Fig.~\ref{fig:Fig_gnd_dn_lambda}. In particular, the unfolded configuration has D/N boundary conditions at $x = \pm \frac{L}{2}$ and linking boundary condition at $x = 0$. The general solutions can be written as
\begin{equation}
\label{eq:normalized_f}
f(k, x) = 
\left\lbrace
\begin{aligned}
  A_1 e^{i kt} \cos\left(kx +\frac{1}{2}kL \right) &  \quad x < 0  \\
  A_2 e^{ikt}  \sin\left(kx - \frac{1}{2}kL \right) & \quad x > 0   \\
\end{aligned} \right. 
\end{equation}
As demonstrated in the \todo[inline]{main text}, if we denote $f(k,x<0)\equiv\phi_1$ and $f(k,x>0)\equiv\phi_2$, the boundary condition at the junction reads
\begin{eqnarray}\begin{aligned}
\frac{\partial_x \phi_1}{ \partial_t \phi_1} = \lambda^2 \frac{\partial_x \phi_2}{ \partial_t \phi_2} = \tan^2 \theta\frac{\partial_x \phi_2}{ \partial_t \phi_2} \quad \theta \in \left[0,\frac{\pi}{2} \right]  
\end{aligned}\end{eqnarray}
which implies
\begin{equation}
\label{eq:momentum_gnd_dn_lambda}
k = \frac{2\pi}{L}\left( n \pm \frac{\theta}{\pi} \right)  \quad n\in\mathbb{Z}
\end{equation}
It is evident that the momentum $k$ is shifted from integer multiple of $2\pi/L$ due to the $\lambda$ boundary condition in the middle. 
\begin{comment}
Thus, the normalized eigenfunctions are
\begin{equation}
f_n(x) = \sqrt{\frac{2}{L}}
\left\lbrace
\begin{aligned}
  \cos(kx +\frac{1}{2}kL ) &  \quad x < 0  \\
  \pm \sin(kx - \frac{1}{2}kL ) & \quad x > 0   \\
\end{aligned} \right. 
\qquad 
k = \frac{2\pi}{L}\left( n \pm  \frac{\theta}{\pi} \right)  \quad n \in \mathbb{Z} 
\end{equation}
Expand the field $\phi = \sum_n \phi_n f_n(x) $, the action and Hamiltonian becomes
\begin{equation}
  S = \frac{g}{2} \int dt \, \sum_{n \in \mathbb{Z} }\left(  \dot{\phi}^2_n + k^2 \phi_n^2 \right) \implies\quad   g \dot{\phi}_n  = \pi_n \quad \implies H =  
\frac{1}{2g}\sum_{n \in \mathbb{Z} } \pi_n^2 + ( kg )^2  \phi_n^2 
\end{equation}
\end{comment}

It is clear that the normalized eigenfunctions in Eq.~\eqref{eq:normalized_f} serves as orthonormal basis in mode expansion. Following the same procedure in \onlinecite{di_francesco_conformal_1997}, we have the Hamiltonian as
\begin{equation}
\label{eq:H_in_gnd_dn_lambda}
H = \frac{1}{2} \sum_{n \in \mathbb{Z} } |k|  \left(a^{\dagger}_n a_n + \frac{1}{2} \right)
\end{equation}
where the momentum $k$ is defined in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, and the creation and annihilation operators defined as usual
\begin{equation}
\begin{aligned}
a_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n + \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
a^{\dagger}_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n - \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
\end{aligned}
\end{equation}

>>>>

In this appendix, we would like to calculate the amplitude for the setup shown in Fig.~\ref{fig:Fig_gnd_dn_lambda}. In particular, the unfolded configuration has D/N boundary conditions at $x = \pm \frac{L}{2}$ and linking boundary condition at $x = 0$. The general solutions can be written as
\begin{equation}
\label{eq:normalized_f}
f(k, x) = 
\left\lbrace
\begin{aligned}
  A_1 e^{i kt} \cos\left(kx +\frac{1}{2}kL \right) &  \quad x < 0  \\
  A_2 e^{ikt}  \sin\left(kx - \frac{1}{2}kL \right) & \quad x > 0   \\
\end{aligned} \right. 
\end{equation}
As demonstrated in Sec.~\ref{sec_sub:general_formulation}, if we denote $f(k,x<0)\equiv\phi_1$ and $f(k,x>0)\equiv\phi_2$, the boundary condition at the junction reads
\begin{eqnarray}\begin{aligned}
\frac{\partial_x \phi_1}{ \partial_t \phi_1} = \lambda^2 \frac{\partial_x \phi_2}{ \partial_t \phi_2} = \tan^2 \theta\frac{\partial_x \phi_2}{ \partial_t \phi_2}, \quad \theta \in \left[0,\frac{\pi}{2} \right]  
\end{aligned}\end{eqnarray}
which implies
\begin{equation}
\label{eq:momentum_gnd_dn_lambda}
k = \frac{2\pi}{L}\left( n \pm \frac{\theta}{\pi} \right),  \quad n\in\mathbb{Z}
\end{equation}
It is evident that the momentum $k$ is shifted from integer multiple of $2\pi/L$ due to the $\lambda$ boundary condition in the middle. 
\begin{comment}
Thus, the normalized eigenfunctions are
\begin{equation}
f_n(x) = \sqrt{\frac{2}{L}}
\left\lbrace
\begin{aligned}
  \cos(kx +\frac{1}{2}kL ) &  \quad x < 0  \\
  \pm \sin(kx - \frac{1}{2}kL ) & \quad x > 0   \\
\end{aligned} \right. 
\qquad 
k = \frac{2\pi}{L}\left( n \pm  \frac{\theta}{\pi} \right)  \quad n \in \mathbb{Z} 
\end{equation}
Expand the field $\phi = \sum_n \phi_n f_n(x) $, the action and Hamiltonian becomes
\begin{equation}
  S = \frac{g}{2} \int dt \, \sum_{n \in \mathbb{Z} }\left(  \dot{\phi}^2_n + k^2 \phi_n^2 \right) \implies\quad   g \dot{\phi}_n  = \pi_n \quad \implies H =  
\frac{1}{2g}\sum_{n \in \mathbb{Z} } \pi_n^2 + ( kg )^2  \phi_n^2 
\end{equation}
\end{comment}

It is clear that the normalized eigenfunctions in Eq.~\eqref{eq:normalized_f} serves as orthonormal basis in mode expansion. Following the same procedure in Ref.~\onlinecite{di_francesco_conformal_1997}, we have the Hamiltonian as
\begin{equation}
\label{eq:H_in_gnd_dn_lambda}
H = \frac{1}{2} \sum_{n \in \mathbb{Z} } |k|  \left(a^{\dagger}_n a_n + \frac{1}{2} \right)
\end{equation}
where the momentum $k$ is defined in Eq.~\eqref{eq:momentum_gnd_dn_lambda}, and the creation and annihilation operators defined as usual
\begin{equation}
\begin{aligned}
a_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n + \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
a^{\dagger}_n = \frac{1}{\sqrt{2}} \left( \sqrt{ |k|g} \phi_n - \frac{i }{\sqrt{|k|g} }\pi_n  \right) \\
\end{aligned}
\end{equation}
